---
title: "Workflow Upgrade: Adverse Events"
output: 
  html_document:
    theme: journal
    toc: true
---

# Adverse Events

Our goal is to explore adverse events for different brand name drugs. To start, we'll access data using the openFDA API to access to information on adverse events.

## Accessing Data

```{r setup, include=FALSE}
# devtools::install_github("ropenhealth/openfda")
library(openfda)
library(dplyr)
library(DBI)
library(ggplot2)
library(ggthemes)
```


Next, we'll write some helper functions to use the openFDA API and package to query adverse events data.

```{r openfda, include=FALSE}
# Helper Functions
get_adverse <- function(gender, brand_name, age) {
  fda_query("/drug/event.json") %>%
    fda_filter("patient.drug.openfda.brand_name", brand_name) %>% 
    fda_filter("patient.patientsex", gender) %>% 
    fda_filter("patient.patientonsetage", age) %>% 
    fda_count("patient.reaction.reactionmeddrapt.exact") %>% 
    fda_limit(10) %>% 
    fda_exec()
}


create_age <- function(min, max){#
  sprintf('[%d+TO+%d]', min, max)
}

```

Using our openFDA helper functions we can pull adverse event data by gender for a specific drug and age range.

```{r pull-data, message=FALSE, warning=FALSE}
sel_name <- 'Tylenol'
age <- create_age(20,60)

male <- get_adverse("1", sel_name,age) %>% 
    mutate(gender = 'male')

female <- get_adverse("2", sel_name, age) %>% 
    mutate(gender = 'female')

adverse <- rbind(male, female)

adverse
```

## Plotting

Now we're ready to start investigating the data and showing off the resulting graphics. We'll create plots and then a static dashboard. 

```{r plots}
adverse %>% 
  group_by(term) %>% 
  summarise(count = sum(count)) %>% 
  ggplot() +
    geom_bar(aes(reorder(term, count), count), stat = "identity") +
    coord_flip() +
    theme_fivethirtyeight() +
    labs(
      title = sel_name
    )
```

```{r gender}
  ggplot(adverse) +
    geom_bar(aes(reorder(term, count), count, fill = gender), stat = "identity") +
    facet_wrap(~gender) +
    coord_flip() +
    theme_fivethirtyeight() +
    labs(
      title = sel_name
    )
```

